Future cosmic microwave background delensing with galaxy surveys

### 2017-12-05 Notes/referee_response.tex

\begin{enumerate}
\item The abstract is unnecessarily very long and could be considerably
shortened.

\vspace{0.4cm}
I agree.
The abstract length is now reduced to $\sim 80\%$ of its previous length, mainly removing redundancy and quantitative results that can be found easily in the tables while leaving the main general points. I believe it is still a little bit longer than the average but reasonably so.
\vspace{1cm}

\item Putting the lmin-lmax ranges in the tables as well as having the
dashed grey line of Fig. 5 onto Fig 6-7-8 would help the readability
of the paper. 
\vspace{0.4cm}

I agree with the referee on this point. I first thought this could cause a bit of clatter, but the gain in readability is worth the risk.
I added the dashed line to all the figures. The $\ell$ range used for the inflationary constraints is also at the end of each table both for readability and reproducibility. The choice was quite arbitrary tough. We used all the scales available to the instrument. The range where the bulk of the signal to noise is, has been explored in some of the referenced work (\cite{dodelson:2014,abazajian:2016} etc.) but not here.

@@

\vspace{1cm}
\item Twice in the paper the author writes that lensing is only one of
secondaries producing B-modes, or his writing gives that impression.
It is unclear to me what is meant by that, as far as I know, galactic
foregrounds should be basically the only expected source of worries
for the experiments considered there.

\vspace{0.4cm}

We agree that the phrasing of the sentences where secondary B-modes sources are introduced can be confusing. While it is well known in the field, it is important to be clear to a broader audience.
For this reason, we add this sentence in the introduction:

\vspace{1cm}
\textit{Two main effects produce non-primordial B-modes: the polarized foregrounds from the Galaxy  and the effect of gravitational interactions with large scale structures (LSS). 
These two components need to be treated with very different techniques and in this work, we will focus on the latter (see \cite{remazeilles:2017b} for a review on the former).}

\vspace{0.2 cm}

Note that if the referee meant that the galactic B-modes component is the only relevant secondary component compared to lensing B-modes, I do not think that is necessarily true. 
While on most of the sky the galactic component is the higher contribution in power, that is not true for an opportunely chose small say 1\% patch. Furthermore, that is probably not the best metric.
As shown in the CMB-S4 science book \cite{abazajian:2016} if one wants to maximize the $r$ constraints, lensing and foregrounds are equally important. 
The importance there is measured as the number of detectors used to either improve the lensing reconstruction or to observe the sky at different frequencies to clean the foregrounds component.
This depends slightly on the fraction of the sky used, but it is true for a wide range that contains all possible future ground experiments.

@@

\vspace{1cm}
\item The author introduces sharp cuts to CIB and WISE at $\ell < 100$. Is the
alternative approach of incorporating the foreground content in the
auto spectra of the tracer (E.g. Eq. 17) combination possible ? That
would retain some more of the signal presumably. 
\vspace{0.4cm}

This is indeed a good idea. The real obstacle here is our current knowledge of foreground spectra. There is not an accurate model for that (at higher scales we use a simple power law). However one might measure that empirically, given a dataset with enough signal to noise on those large scales, by measuring $C^{II}$ and $C^{I\phi}$ directly from the data. Doing that will call everything in the tracer map that does not correlate with $\phi$, "noise". That noise will include foregrounds too. However, it is unclear if eventual dust residual can cause biases in the delensed B-modes (if they correlate with dust contamination in CMB polarizations).
Given the small amount of delensing power for $\ell<100$ the field decided, so far, to cut those in forecasts. Given how much foregrounds are complicated I believed only data would answer this. An exploration of these on simulations is probably beyond the scope of this paper, but very soon future delensing work will test this on real data.

@@

\vspace{1cm}
\item Fig 1 might look confusing at first. The SKA kernel there shows
brilliant overlap with the CMB lensing kernel, but is not the most
powerful tracer. On the other hand, DES for instance has very moderate
overlap but can still achieve very high performance on large scales.
Can the author comment and help the reader make sense of that figure ?

\vspace{0.4cm}
I agree that Fig 1 may be a little confusing because it does not tell the entire story. 
The redshift overlap is not the only ingredient for the delensing efficiency. For example, the noise is relevant as well, and this is why for example for S2 experiments CIB can be better than the CMB internal reconstruction even if the latter has a perfect overlap. 
Additionally, not all the redshifts matter at the same level since they correspond to different angular scales (depending on the distance of the tracer) which, as shown by the dashed curves in the figures 5-8, contributes differently to the generation of CMB B-modes.
Showing all of this in one plot is hard.

@@

However, I think Fig 1 tells most of the story. For example, the referee is right that DES is less efficient than SKA in delensing efficiency (16\%) vs (52\%).
Furthermore, I think this figure highlights the central part (the "kernel") that characterize different tracers spectra (using Eq. 2) for people not familiar with it.

I added to the caption of Fig. 1 a sentence to highlight that the efficiency is not proportional to the overlapping area of this plot because different redshifts contribute differently due to instrumental noise and the geometric properties of the lensing kernel.
\vspace{1cm}

@@

\item Is it clear, as the author claims in IV A with no supporting
evidence, that fixing the cosmology has no impact on the {r,nt}
constraints ? The cosmology fixes the residual B-mode amplitude, which
is unequally important in the delensed vs nominal cases. An slightly
improved discussion would be welcome here.

\vspace{0.4cm}
Note that indeed, as specified in the text, the chosen cosmology \textit{has} an impact on the absolute value of the ($r$,$n_{t}$) constraints.
What I claim cosmology has almost no impact on, in particular if compared to other uncertainties, is the delensing efficiency (how much these constraints improve), i.e. the ratio $C^{BB}_{\ell}$ nominal/delensed. What matters for the efficiency is the \textit{fraction} of removed power. That depends only on the factor  $1-\rho^2$ where $\rho$ is the correlation factor. 
The correlation factor has a very little dependence on cosmology in particular given the tight constraints from Planck on $\Omega_{m}h^2$ and $\sigma_{8}$, and it is expected to be a secondary uncertainty here compared to, for example, the redshift distribution of tracers. Indeed $\rho$ contains $C^{II}$ $C^{I\phi}$, $C^{\phi\phi}$ that are integrals, with slightly different weights in redshift, of the cosmology dependent power spectrum $P(k)$. For this reason, they should scale similarly with cosmology leading to a substantial cancellation in the cosmological dependence of $\rho$.

Different cosmologies can also shifts the scales where the information on ($r$, $n_{t}$) come from by changing the shape of the $C_{\ell}^{BB}$ spectra.

I agree that testing this intuition is important. For this reason, I run the same pipeline with $10\%$ higher $\Omega h^{2}$, which is almost 40 times more than what Planck constraints \citep{planck-collaboration:2016b}. The efficiency values changes, but by at most $8-9\%$. Given that the Planck constraints are way tighter I consider this, as expected, a present but subdominant effect. 
For example for $1\%$ higher $\Omega h^{2}$ the efficiency values changes by at most $1.5\%$
I added few sentences to the paper to report this test.

@@

\vspace{1cm}

\item The author makes a strong point that survey data should be split
into bins, as this makes the correlation increases by 30\%. I am not
entirely convinced that this is as relevant as claimed. From Fig 4, it
seems that the improvement is fairly limited on the scales that are
relevant for B-mode delensing. Presumably, that is more important for
T or E delensing. The author might want to comment or adapt.

\vspace{0.4cm}
In general, I agree with the point that quoting the improvement in the correlation is not entirely fair because the scales at which that improvement happens is also necessary to understand how much power will be removed.
For this reason, I added in the abstract this quantity too.

Furthermore, while I agree that the improvement would be even more significant for T-E delensing, it is nonetheless also quite important for BB too.
As can be seen in Figure 4 in the paper LSST shows a significant improvement in the correlation with the lensing potential all the way to $\ell\sim 800$. Even for other probes, in general, less powerful, there is still an improvement for example at \textit{all} the scales where the contribution to BB (the dashed line) is relevant $50<\ell<400$.
Below I am attaching a table that shows the improvement in the BB power removed and constraints on $\sigma(r)$.
Note that the 30\% level quoted in the abstract corresponds to the most futuristic tracer, i.e., LSST. That should be explicitly stated, so I added this to the abstract.

@@

\item
The author uses iterative internal delensing noise levels for the
CMB. I believe this could be discussed in more details. Can he specify
what multipole range is considered for this calculation ? I suppose
this multipole range should not overlap with the primordial B that is
used for constraints on r, however, this range (unnecessarily) extends
up to 800 in the S3 section for example. Most likely, 200 would be
largely enough, does it imply a sizable reduction on the iterative
reconstruction noise levels ? Besides, iterative delensing is
complicated and not established yet, in contrast to Quadratic
Estimator delensing. The paper might gain in interest if numbers for
Q.E. and iterative delensing were presented.

\vspace{0.4cm}

%One information that is probably not clear enough or partially missing in the text is that iterative delensing compared to iterative delensing is important only for CMB-S4 noise levels. 
%For example, the amount of BB power removed for CMB-S3 level using iterative vs. QE is 51\% vs. 56\%.

Let me start from the first part of the comment which regards the scales used here. 
For reproducibility, I added the scales used for the lensing reconstruction for the different CMB stages.
Furthermore, the idea of a separation in scale between the CMB data used to reconstruct the potential and the data used to constrained $r$, is indeed an interesting idea proposed (for example in \cite{sehgal:2017}) to solve the so-called "internal delensing bias". This is important for both QE and iterative approach. This and other methods have shown that, at least on simulations, we can avoid this bias \textit{without losing performance}. 
Other groups have shown that it is possible to use overlapping scales and then correct for the bias.
For this reason, we assume here that we can reconstruct the lensing potential using scales that overlap with the primordial B-mode, without being over-optimistic. 
While a full treatment of this effects with realistic simulations is important, we left a full study for future works.
Again this can also be tested on data. For example, the deep 100 deg$^{2}$ field measured by SPTPol that can reach a depth of $5-6 ~\mu$K arcmin with high resolution might be the the first dataset where this can be explored. 
I comment on this at the end of section IIA.

@@

Regarding the second part of the comment, I agree that it is interesting to quickly investigate the loss in performance if the quadratic estimator is used instead of the iterative one. For this reason, I added a new case for S3 and S4 experiment: the quadratic estimator case. Since this is a widely used technique, I think it can be presented without much introduction. I added a few sentences at the end of section IIA and in the S3 and S4 sections. Note that there is no difference between these two approaches for current levels of noise (S2).

%Following the referee comments, we
%\begin{itemize}
%\item Make it clear that only CMB S4 will benefit from iterative delensing.
%\item Add the level of delensing that is possible to achieve with CMB-S4 if the well tested QE is used instead of the iterative approach. This gives in some sense a range of possible efficiency for CMB S4 experiments.
%\end{itemize}
%
% 
\end{enumerate}

### 2017-12-05 Notes/galaxy_delensing.tex
major working on referee response

The primordial B-modes, a particular component of the cosmic microwave background (CMB) polarization are a promising experimental dataset to probe the inflationary paradigm. B-modes are indeed a direct consequence of the presence of gravitational waves in the early universe. However, several secondary effects in the low redshift universe will produce \textit{non-primordial} B-modes. In particular, the gravitational interactions of CMB photons with large-scale structures will distort the primordial E-modes, adding a lensing B-mode component to the primordial signal. Removing the lensing part ("delensing") from the measurement of CMB B-modes will then be necessary to constrain the amplitude of the primordial gravitational waves. Here we discuss the role of current and future large-scale structure surveys in a multi-tracers approach to CMB delensing. We find that, in general, galaxy surveys should be split into tomographic bins as this can improve the reconstruction of the lensing potential that lenses the CMB photons by 30\%. 
Ongoing or recently completed CMB experiments (CMB-S2) will particularly benefit from large-scale structure tracers that, once properly combined, will have a better performance than a CMB internal reconstruction. With the decrease of instrumental noise, the CMB lensing internal reconstruction will improve, and the fraction of removed lensing B-modes using CMB alone will rapidly improve for CMB S3 and S4. Nevertheless, optical galaxy surveys will still play an important role even for CMB S4. In particular, an LSST-like survey can a achieve a delensing performance comparable to a 3G CMB experiment but with entirely different systematics. This redundancy will be essential to prove the robustness against systematics of an eventual detection of primordial B-modes.

@@

Because the CMB B-modes provide the cleanest known observational window into the primordial gravitational wave background, improving their measurement is a major objective of current and future CMB experiments.
Even if inflationary B-modes have not been detected yet, the target value $r \gtrsim 10^{-3}$, should be reachable shortly given the level of noise expected in future CMB experiments \citep{kamionkowski:2016,abazajian:2016}.
However, just reducing the level of noise will not be enough to reach this target.
Indeed, the observed B modes are not solely sourced by early universe physics; they are also produced by secondary effects taking place in the late-time low-redshift universe.
Two main effects produce non-primordial B-modes: the polarized foregrounds from our own galaxy  and the effect of gravitational interactions with large scale structures (LSS). 
This two components needs to be treated with very different techniques and in this work we will focus on the latter (see \cite{remazeilles:2017b} for a review on the former).
Lensing shears the CMB polarization pattern, producing ``lensing B modes'' from CMB E modes \citep{zaldarriaga98}.
This expected component has now been measured both in cross-correlation with LSS \citep{hanson13,bicep2-collaboration:2016,polarbear2014c,van-engelen:2015,planck2015XV} and from CMB data alone \citep{polarbear2014b,bicep2a,keisler15}.

@@

The CMB lensing potential is \textit{the} field that we need in order to reverse the effect of large scale structure and delens the CMB.
However, the lensing potential can also be reconstructed using the CMB itself. In that case, we can treat it as a noisy tracer of the true field. Both the CMB lensing field and its noisy reconstructed counterpart have the same kernel $W^{\kappa}(z)$.
However, when computing the power spectrum of the latter, we need to add a noise component. Given the instrumental noise level and the beam, we can calculate the reconstruction noise $N_{\ell}^{\kappa \kappa}$, and so
\beq\label{eqn:noisekappa}
\cl^{\kappa_\mathrm{rec} \kappa_\mathrm{rec}} = \cl^{\kappa \kappa } +N^{\kappa \kappa}_{\ell}
\eeq
In this work, the level of noise is computed assuming an iterative approach to the CMB lensing reconstruction as described in \cite{smith:2012,hirata:2003}. The iterative approach has not been fully demonstrated on data yet but it is among the major goals of the CMB community (\cite{abazajian:2016}).
Note however that, as pointed out in in \cite{smith:2012,hirata:2003}, this improves the CMB lensing reconstruction compared to a quadratic estimator approach mainly for CMB-S4 noise levels.

@@

% ===================
\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=1.]{./images/compare_kernel.pdf}
\caption{Kernel Comparison: Comparison of the different kernels as a function of redshift for some of the tracers used in this analysis. The larger the overlap with the CMB lensing kernel the better the reconstruction of the lensing potential is thus leading to a higher delensing efficiency. 
However, the delensing efficiency is not exactly proportional to the area of overlap here. Indeed each redshift contributes differently to CMB lensing because of both the noise in the measurement and the geometric properties of the lensing kernel.
Note that for galaxy surveys the kernel corresponds to $b(z)\frac{dN}{dz}$ and not just the redshift distribution of galaxies.}
\label{fig:kernel}
\end{center}
\end{figure}
% ===================

@@

In \refeq{fisher} we are making a few important assumptions. 
First, we are fixing all the cosmological parameters except $\{r,n_{t}\}$. Uncertainties in those will propagate to larger uncertainties in $\{r,n_{t}\}$. While neglecting this will lead to slightly optimistic absolute constraints, it has no significant impact on the estimate of the improvement due to delensing. For example, running the same pipeline with $10\%$ higher $\Omega h^{2}$, which is almost 40 times more than what Planck constraints \cite{planck-collaboration:2016b}, change the delensing improvement factors by at most $6-7\%$.

@@

On the CMB side, we will combine a BICEP-Keck like \textit{deep} CMB experiment with an overlapping \textit{higher resolution} experiment. For the deep experiment we assume an instrumental noise equal to $3\mu$K-arcmin in polarization ($\sqrt{2}$ lower in temperature), a beam of 30 arcmin and an angular scale range of $50<\ell<500$.
We assume that the CMB lensing reconstruction is performed by the higher resolution CMB experiment.
We explore two possibilities.
First, we test the delensing efficiency with an internal reconstruction of CMB lensing performed by the Planck satellite \cite{planck-collaboration:2016a}. In this case, for the noise in the CMB lensing map in \refeq{noisekappa} we use the actual noise curves publicly available
\footnote{\url{https://wiki.cosmos.esa.int/planckpla2015/index.php/Specially\_processed\_maps}}. For consistency, we also combine a Planck-like E-mode map of $60\mu$K-arcmin in polarization with a beam equal to 7 arcmin that however adds almost nothing to the \textit{deep} CMB experiment.
Then, we test a second scenario with a ground-based experiment characterized by noise levels in polarization consistent with SPT-Pol \cite{henning:2017}: $9.4\mu$K-arcmin in polarization with a beam equal to 1.2 arcmin. In this case we set the angular scale range at $50<\ell<4000$. However, the results are quite robust against this choice.

We combine these CMB experiments with the CIB and current low redshift galaxy surveys like DES and WISE and we compute the improvement in the delensing efficiency.
Following \cite{yu:2017} we cut both the CIB and WISE at $\ell<100$ where they are contaminated by large Galactic dust residuals.
Optical surveys galaxies like DES are less affected by dust and can be used on larger scales.
The achievable correlation is shown in \reffig{corrnow-planck} for Planck and in \reffig{corrnow} for SPT-Pol.
In these figures, in order to show which scales in $\kappa$ contribute to the $\ell<100$ B-mode power, we include a dashed curve that corresponds to (with arbitrary scale) $<C_{\ell}^{\kappa \kappa} \times \frac{\partial C^{BB}_{\ell'}}{\partial{C^{\kappa \kappa}_{\ell}}}>_{\ell ' <100}$.

@@

\subsection{CMB-S3 Era}\label{sec:S3}
CMB polarization measurements are rapidly improving. Indeed the next generation of ground-based telescopes has been already deployed, and data are currently being taken.
As we did for Stage-2, we model the CMB S3 as two different overlapping experiments.
For the deep experiment we assume an instrumental noise equal to $2\mu$K-arcmin in polarization and a beam of 30 arcmin.
The high resolution experiment will have a level of noise in polarization of $3\mu$K-arcmin and a 1 arcmin beam. This level of noise is also assumed for the internal noise reconstruction.
We use an angular scale range in the Fisher matrix of $50<\ell<800$ even if most of the high $\ell$ scale do not contribute to the constraints.
Furthermore, around 2019, DESI will start taking data. For this reason, we add DESI to the LSS tracers used in \refsec{del_res}.

The correlation factor attainable using generation 3 experiments is shown in \reffig{corrS3}.
An interesting finding is that DESI will be less efficient (removing $11\%$ of the power) than a DES-like survey ($17\%$) despite the fact that it can probe slightly higher redshift. The reason is that, because of the broad CMB kernel, spectroscopic redshift accuracy is not needed and the lower shot noise in DES increases the delensing efficiency. Unfortunately, adding DESI will only bring the power removed using LSS from 41 to 45\%.
Contrary to current S2 experiments, the CMB internal reconstruction will dominate the correlation with the lensing potential up to $\ell\simeq 600$.
For CMB S3, the improvement due to an iterative approach compared to the quadratic estimator is still minimal so that the more traditional technique can be used. For example, in this case, the removed power goes from $56\%$ to $54\%$ (quoted in \reftab{S3}) when using a quadratic estimator instead of the iterative one.

@@

\subsection{CMB-S4 Era}\label{sec:S4}
Having as a major goal the detection of inflationary B-modes, CMB data will continue improving even after Stage-3.
An ambitious program for a Stage-4 ground CMB experiment is in the planning phase \cite{abazajian:2016}. 
Moreover, satellite and balloon CMB experiments have been proposed and have the potential to extend the accessible B-mode measurements to the largest scales. Given the unprecedented low level of noise of these experiment, delensing will be even more important than in previous generations.
Following the B-mode constraints section in \cite{abazajian:2016}, here we assume a CMB-S4 ground experiment composed of two different telescopes.
For the deep experiment, we assume an instrumental noise equal to $1\mu$K-arcmin in polarization and a beam of 15 arcmin. The high-resolution experiment will have a level of noise in polarization of $1.5\mu$K-arcmin and a 1 arcmin beam.

@@

\reftab{S4} shows the improvement on the inflationary constraint given the cross-correlation factors described above. 
LSST alone will be able to remove almost half of the B-mode lensing power. This will lead to an improvement in $\sigma(r)$ of a factor of 2. If we add all the other infrared and optical LSS tracers previously included (i.e. excluding SKA) the improvement factor increase to 2.4.
These gives an important opportunity: for sufficiently high values of $r$ it will be possible to confirm a detection of primordial gravitational waves using spectra delensed with LSS, completely independent of CMB internal data. 
Similarly to \cite{namikawa:2016a} we find that SKA phase-1 radio survey will remove 53\% of the lensing power and, once combined with optical and infrared surveys, this will allow the level of power removed using LSS (62\%) to surpass the level of the CMB S3 internal reconstruction.
Finally, as previously mentioned, CMB S4 will be the main source of delensing with the ability of removing up to more than 80\% of the B-modes power. This will bring a factor of $\sim 4$ improvement in the null test case even with realistic noise. Combining with LSS will only increase the removed power compared to CMB only by $\simeq 4.3\%$ but it will be highly beneficial to test the robustness of the final result.

>>>>

The primordial B-modes component of the cosmic microwave background (CMB) polarization is a promising experimental dataset to probe the inflationary paradigm. B-modes are indeed a direct consequence of the presence of gravitational waves in the early universe. However, several secondary effects in the low redshift universe will produce \textit{non-primordial} B-modes. In particular, the gravitational interactions of CMB photons with large-scale structures will distort the primordial E-modes, adding a lensing B-mode component to the primordial signal. Removing the lensing component ("delensing") will then be necessary to constrain the amplitude of the primordial gravitational waves. Here we discuss the role of current and future large-scale structure surveys in a multi-tracers approach to CMB delensing. We find that, in general, galaxy surveys should be split into tomographic bins as this can increase the reduction of lensing B-modes by $\sim 25\%$ in power in the most futuristic case. 
Ongoing or recently completed CMB experiments (CMB-S2) will particularly benefit from large-scale structure tracers that, once properly combined, will have a better performance than a CMB internal reconstruction. With the decrease of instrumental noise, the amount of lensing B-modes power removed using CMB internal reconstruction alone will rapidly increase. Nevertheless, optical galaxy surveys will still play an important role even for CMB S4. In particular, an LSST-like survey can a achieve a delensing performance comparable to a 3G CMB experiment but with entirely different systematics. This redundancy will be essential to prove the robustness against systematics of an eventual detection of primordial B-modes.

@@

Because the CMB B-modes provide the cleanest known observational window into the primordial gravitational wave background, improving their measurement is a major objective of current and future CMB experiments.
Even if inflationary B-modes have not been detected yet, the target value $r \gtrsim 10^{-3}$, should be reachable shortly given the level of noise expected in future CMB experiments \citep{kamionkowski:2016,abazajian:2016}.
However, just reducing the level of noise will not be enough to reach this target.
Indeed, the observed B modes are not solely sourced by early universe physics; they are also produced by secondary effects taking place in the late-time low-redshift universe.
\textcolor{blue}{Two main effects produce non-primordial B-modes: the polarized foregrounds from the Galaxy  and the effect of gravitational interactions with large scale structures (LSS). 
These two components need to be treated with very different techniques, and in this work, we will focus on the latter (see \cite{remazeilles:2017b} for a review on the former).}
Lensing shears the CMB polarization pattern, producing ``lensing B modes'' from CMB E modes \citep{zaldarriaga98}.
This expected component has now been measured both in cross-correlation with LSS \citep{hanson13,bicep2-collaboration:2016,polarbear2014c,van-engelen:2015,planck2015XV} and from CMB data alone \citep{polarbear2014b,bicep2a,keisler15}.

@@

The CMB lensing potential is \textit{the} field that we need in order to reverse the effect of large scale structure and delens the CMB.
However, the lensing potential can also be reconstructed using the CMB itself. In that case, we can treat it as a noisy tracer of the true field. Both the CMB lensing field and its noisy reconstructed counterpart have the same kernel $W^{\kappa}(z)$.
However, when computing the power spectrum of the latter, we need to add a noise component. Given the instrumental noise level and the beam, we can calculate the reconstruction noise $N_{\ell}^{\kappa \kappa}$, and so
\beq\label{eqn:noisekappa}
\cl^{\kappa_\mathrm{rec} \kappa_\mathrm{rec}} = \cl^{\kappa \kappa } +N^{\kappa \kappa}_{\ell}
\eeq
In this work, the level of noise is computed assuming an iterative approach to the CMB lensing reconstruction as described in \cite{smith:2012,hirata:2003}. \textcolor{blue}{The iterative approach has not been sufficiently demonstrated on data yet, but it is among the future goals of the CMB community \cite{abazajian:2016}.
Note however that, as pointed out in \cite{smith:2012,hirata:2003}, this improves the CMB lensing reconstruction compared to a quadratic estimator approach mainly for S3 and S4 noise levels. For these two cases, we also show the delensing efficiency obtained when the CMB lensing is reconstructed with the standard quadratic estimator (QE) approach \cite{okamoto:2003,hu:2002}.
Furthermore, we do not deal here with the presence of an internal delensing bias. Several promising approaches have been proposed \cite{sehgal:2017,namikawa:2017} even if they have not been applied to low noise data yet \cite{carron:2017}.}

@@

% ===================
\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=1.]{./images/compare_kernel.pdf}
\caption{Kernel Comparison: Comparison of the different kernels as a function of redshift for some of the tracers used in this analysis. The larger the overlap with the CMB lensing kernel the better the reconstruction of the lensing potential is thus leading to a higher delensing efficiency. 
\textcolor{blue}{However, the efficiency is not exactly proportional to the overlapping area. 
Firstly structure at different redshifts contribute differently to the generation of large-scale B-modes because of the geometric properties of the lensing kernel, and secondly, we are not taking the tracers noise into account here.
Note that for galaxy surveys the kernel corresponds to $b(z)\frac{dN}{dz}$ and not just the redshift distribution of galaxies.}}
\label{fig:kernel}
\end{center}
\end{figure}
% ===================

@@

In \refeq{fisher} we are making a few important assumptions. 
First, we are fixing all the cosmological parameters except $\{r,n_{t}\}$. Uncertainties in those will propagate to larger uncertainties in $\{r,n_{t}\}$. While neglecting this will lead to slightly optimistic absolute constraints, it has no significant impact on the estimate of the improvement due to delensing. \textcolor{blue}{For example, running the same pipeline with $10\%$ higher $\Omega h^{2}$, which is almost 40 times more than what Planck constraints \cite{planck-collaboration:2016b}, change the delensing improvement factors by at most $7-8\%$. A more reasonable 1\% higher $\Omega h^{2}$ corresponds to variations in the improvement factors of at most 1.5\%.}

@@

On the CMB side, we will combine a BICEP-Keck like \textit{deep} CMB experiment with an overlapping \textit{higher resolution} experiment. For the deep experiment we assume an instrumental noise equal to $3\mu$K-arcmin in polarization ($\sqrt{2}$ lower in temperature), a beam of 30 arcmin and an angular scale range of $50<\ell<500$.
\textcolor{blue}{Here and in the following sections, we use all scales available to the instrument, even if the bulk of the signal to noise is at large scales $\ell<500$ \cite{dodelson:2014}.}
We assume that the CMB lensing reconstruction is performed by the higher resolution CMB experiment.
We explore two possibilities.
First, we test the delensing efficiency with an internal reconstruction of CMB lensing performed by the Planck satellite \cite{planck-collaboration:2016a}. In this case, for the noise in the CMB lensing map in \refeq{noisekappa} we use the actual noise curves publicly available
\footnote{\url{https://wiki.cosmos.esa.int/planckpla2015/index.php/Specially\_processed\_maps}}. For consistency, we also combine a Planck-like E-mode map of $60\mu$K-arcmin in polarization with a beam equal to 7 arcmin that however adds almost nothing to the \textit{deep} CMB experiment.
Then, we test a second scenario with a ground-based experiment characterized by noise levels in polarization consistent with SPT-Pol \cite{henning:2017}: $9.4\mu$K-arcmin in polarization with a beam equal to 1.2 arcmin. In this case we set the angular scale range at $50<\ell<3000$. However, the results are quite robust against this choice.

We combine these CMB experiments with the CIB and current low redshift galaxy surveys like DES and WISE and we compute the improvement in the delensing efficiency.
Following \cite{yu:2017} we cut both the CIB and WISE at $\ell<100$ where they are contaminated by large Galactic dust residuals. 
\textcolor{blue}{In the future, with better theoretical foreground models or high signal to noise measurements, we might be able to characterize these fields (and their spectra $C_{\ell}^{\kappa I}$ and $C_{\ell}^{II}$) even at larger scales. Potential biases effect may be present like, for example, the correlation between the foreground contamination in galaxies and in the CMB polarization itself. 
If these are under control, we might then be able to also use the $\ell<100$ component of these tracers to delens. However since the lensing contribution to B-modes from LSS modes at these scales is quite small, there is not too much to gain in term of delensing efficiency.}

Optical surveys galaxies like DES are less affected by dust and can be used on larger scales.
The achievable correlation is shown in \reffig{corrnow-planck} for Planck and in \reffig{corrnow} for SPT-Pol.
In these figures, in order to show which scales in $\kappa$ contribute to the $\ell<100$ B-mode power, we include a dashed curve that corresponds to (with arbitrary scale) $<C_{\ell}^{\kappa \kappa} \times \frac{\partial C^{BB}_{\ell'}}{\partial{C^{\kappa \kappa}_{\ell}}}>_{\ell ' <100}$.

@@

\subsection{CMB-S3 Era}\label{sec:S3}
CMB polarization measurements are rapidly improving. Indeed the next generation of ground-based telescopes has been already deployed, and data are currently being taken.
As we did for Stage-2, we model the CMB S3 as two different overlapping experiments.
For the deep experiment we assume an instrumental noise equal to $2\mu$K-arcmin in polarization and a beam of 30 arcmin.
The high resolution experiment will have a level of noise in polarization of $3\mu$K-arcmin and a 1 arcmin beam. This level of noise is also assumed for the internal noise reconstruction. For the CMB internal lensing reconstruction we use scales $50<\ell<3000$.
We use an angular scale range in the Fisher matrix of $50<\ell<800$ even if most of the high $\ell$ scale do not contribute to the constraints.
Furthermore, around 2019, DESI will start taking data. For this reason, we add DESI to the LSS tracers used in \refsec{del_res}.

The correlation factor attainable using generation 3 experiments is shown in \reffig{corrS3}.
An interesting finding is that DESI will be less efficient (removing $11\%$ of the power) than a DES-like survey ($17\%$) despite the fact that it can probe slightly higher redshift. The reason is that, because of the broad CMB kernel, spectroscopic redshift accuracy is not needed and the lower shot noise in DES increases the delensing efficiency. Unfortunately, adding DESI will only bring the power removed using LSS from 41 to 45\%.
Contrary to current S2 experiments, the CMB internal reconstruction will dominate the correlation with the lensing potential up to $\ell\simeq 600$.
\textcolor{blue}{For CMB S3, the improvement due to an iterative approach compared to the quadratic estimator is present but not significant, so that a more traditional technique can be used. For example, in this case, the removed power goes from $56\%$ to $51\%$ (quoted in \reftab{S3}) when using a quadratic estimator (QE) instead of the iterative one.}

@@

\subsection{CMB-S4 Era}\label{sec:S4}
Having as a major goal the detection of inflationary B-modes, CMB data will continue improving even after Stage-3.
An ambitious program for a Stage-4 ground CMB experiment is in the planning phase \cite{abazajian:2016}. 
Moreover, satellite and balloon CMB experiments have been proposed and have the potential to extend the accessible B-mode measurements to the largest scales. Given the unprecedented low level of noise of these experiment, delensing will be even more important than in previous generations.
Following the B-mode constraints section in \cite{abazajian:2016}, here we assume a CMB-S4 ground experiment composed of two different telescopes.
For the deep experiment, we assume an instrumental noise equal to $1\mu$K-arcmin in polarization and a beam of 15 arcmin. The high-resolution experiment will have a level of noise in polarization of $1.5\mu$K-arcmin and a 1 arcmin beam. Here we use the same angular scale range in the Fisher as the S3 case and the CMB internal lensing reconstruction is done using scales $50<\ell<4000$.

@@

\reftab{S4} shows the improvement on the inflationary constraint given the cross-correlation factors described above. 
LSST alone will be able to remove almost half of the B-mode lensing power. This will lead to an improvement in $\sigma(r)$ of a factor of 2. If we add all the other infrared and optical LSS tracers previously included (i.e. excluding SKA) the improvement factor increase to 2.4.
These gives an important opportunity: for sufficiently high values of $r$ it will be possible to confirm a detection of primordial gravitational waves using spectra delensed with LSS, completely independent of CMB internal data. 
Similarly to \cite{namikawa:2016a} we find that SKA phase-1 radio survey will remove 53\% of the lensing power and, once combined with optical and infrared surveys, this will allow the level of power removed using LSS (62\%) to surpass the level of the CMB S3 internal reconstruction.
Finally, as previously mentioned, CMB S4 will be the main source of delensing with the ability of removing up to more than 80\% of the B-modes power. 
\textcolor{blue}{Note that, at the noise levels of CMB-S4 it will be quite valuable to develop an iterative algorithm for lensing reconstruction. As shown in \reftab{S4}, a quadratic estimator will only be able to remove $73\%$ of power from $C^{BB}$ compared to the 83\% of the iterative algorithm.}
This will bring a factor of $\sim 4$ improvement in the null test case even with realistic noise. Combining with LSS will only increase the removed power compared to CMB only by $\simeq 4.3\%$ but it will be highly beneficial to test the robustness of the final result.
